---
title: "ggacmap introduction"
author: "Guangchuang Yu"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: false
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{ggacmap}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---


The `ggacmap` function is written by `ggplot2` and take the advantages of using `scale` or `theme` to modify the details.

## Loading the packages

```{r}
library(Racmacs)
library(ggplot2)
library(ggacmap)
```

## Load the H3 AC map

```{r}
f <- system.file("extdata/Day1/files/h3map2004.ace", package="ggacmap")
h3map <- read.acmap(f)
```

## Plotting examples

Plot the map is easy, just using `ggacmap(your_acmap)`:

```{r}
p <- ggacmap(h3map)
print(p)
```

### Change size of points

```{r warning=F, message=F}
p + scale_size(range = 6)
```

### Show color legend

```{r}
p <- p + theme(legend.position='right')
print(p)
```

Oops, this is not what we want. `ggplot2` does the right thing as we map the color to the actual color vector.

How if we want to label these groups with representative strains?

```{r message=F}
## get the first strain of each group
cols <- agFill(h3map)
i <- which(!duplicated(cols))
strain <- rownames(agCoords(h3map))[i]
print(strain)

## a simplify version
sstrain <- sub("/\\d+", "", strain)
print(sstrain)

### Changing the label of color legend
p + scale_color_identity("Antigenic group",
                         labels = sstrain,
                         breaks = cols[i],
                         guide = "legend")
```

### Changing color scheme

```{r message=F}
p <- p + scale_color_viridis_d("Antigenic group",
                          labels = sstrain,
                          breaks = cols[i],
                          guide = "legend")
print(p)
```

### Change point shapes

```{r message=F}
p + scale_shape_manual(values = c(3, 4))
```


### Remove `type` legend

```{r}
p + guides(shape =  FALSE) 
```

## Bonus: easy to put several figures in one page


The following example build a quick-and-dirty tree from the aligned sequences, and visualize the tree using ggtree with tip points colored by antigenic grouping information. We can therefore check whether the antigenic group information is consistent with genetic distance or not.


```{r fig.width=7.5, fig.height=5, warning=F}
seq_file <- system.file("extdata/Day2/files/h3_2004/sequences.csv", package="ggacmap")
sequence <- read.csv(seq_file,
                     colClasses='character',
                     row.names=1,
                     check.names = FALSE,
                     stringsAsFactors=FALSE)

aaseq <- apply(sequence, 1, paste0, collapse='')
aaseq <- Biostrings::AAStringSet(aaseq)

tree <- ape::bionj(Biostrings::stringDist(aaseq, method='hamming'))

d <- data.frame(label = rownames(agCoords(h3map)),
                color = agFill(h3map)
                )

require(ggtree)
p2 <- ggtree(tree) %<+% d + geom_point(aes(color = I(color)))
cowplot::plot_grid(p2, ggacmap(h3map), ncol=2, rel_widths=c(1, .4))
```
